---
title: "Assignment1"
author: "Arshad!"
date: "13/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


1:)   Loading and processing the data

```{r}

rm(list=ls())

library(dplyr)

library(lubridate)

library(ggplot2)

##   Set working Directory

setwd("C:/Users/arsha/OneDrive/Desktop/reproduceable_research/assignment1")

mydata <- read.csv("repdata_data_activity/activity.csv", header=TRUE, na.strings="NA")

##cleanup date class

mydata$date <-ymd(mydata$date)

##Remove NA

mydata1 <- na.omit(mydata)

summary(mydata1)

str(mydata1)

head(mydata1)

tail(mydata1)


```

2:)   Make an histtogram of the total no of steps taken each day

```{r}

mydata2 <- summarize(group_by(mydata1,date),daily.step=sum(steps))

mean.mydata <- as.integer(mean(mydata2$daily.step))

median.mydata <- as.integer(median(mydata2$daily.step))

plot.steps.day <- ggplot(mydata2, aes(x=daily.step)) +
  geom_histogram(binwidth = 600, aes(y=..count.., fill=..count..)) +
  geom_vline(xintercept=mean.mydata, colour = "Black", linetype="dashed", size=1) +
  geom_vline(xintercept=median.mydata, color = "red", linetype="dotted", size=1) +
  labs(title="Number of Steps Taken Each Day", y="Frequency", x="Steps Each Day")

 plot.steps.day
  

```

3:)   Calculate Mean and Median of Total no of steps taken per day...

```{r}


##   Calculating Mean 

mean.mydata

##   Calculating Median

median.mydata

```

4:)   Makeing a time series plot avg no of steps taken

```{r}

mydata3 <- mydata1 %>% group_by(interval) %>% summarize(mean.step=mean(steps))

plot.step.interval <- ggplot(mydata3, aes(x=interval, y=mean.step)) +
  geom_line(color="blue")  +
  labs(title="Avg no of Steps Taken 5-min Interval", y="Avg no of Steps", x="5-min Interval Time Series")

plot.step.interval

```

5:)   Max no of step taken in avg, 5-min Interval

```{r}

anab <- which.max(mydata3$mean.step)

anab.step <- mydata3$interval[anab]

sprintf("Max no of Steps is from %gth 5-min Interval", anab.step)

```

6:)  Planning for Imputing the missing data

```{r}

# Total no of missing values in Dataset.
sum(is.na(mydata))

fas.mydata <- mydata

fas.mydata$steps[is.na(fas.mydata$steps)] <- mean(fas.mydata$steps,na.rm=TRUE)

fas.mydata$steps <- as.numeric(fas.mydata$steps)

fas.mydata$interval <- as.numeric(fas.mydata$interval)

colSums(is.na(fas.mydata))

# Creating a new Dataset with missing values of original Dataset 

summary(fas.mydata)

```

7:)   Histogram of the total number of steps taken each day after missing values are imputed

```{r}

fas.mydata2 <- summarize(group_by(fas.mydata,date), daily.step=sum(steps))

mean.fas <- as.integer(mean(fas.mydata2$daily.step))

mean.fas

median.fas <- as.integer(median(fas.mydata2$daily.step))

median.fas

plot.steps.day <- ggplot(fas.mydata2, aes(x=daily.step)) +
  geom_histogram(binwidth = 600, aes(y=..count.., fill=..count..)) +
  geom_vline(xintercept=mean.fas, color = "Black", linetype="dashed", size=1) +
  geom_vline(xintercept=median.fas, color = "red", linetype="dotted", size=1) +
  labs(title="Number of Steps Taken Each Day", y="Frequency", x="Steps Each Day")

plot.steps.day

```

8:)   Finding the difference in activity patterns between weekdays and weekends

```{r}

fas.mydata$day <- ifelse(weekdays(fas.mydata$date) %in% c("Saturday", "Sunday"), "weekday", "weekend")

fas.df <- fas.mydata %>% group_by(interval,day) %>% summarise(mean.step=mean(steps))

plot.weekday.interval <- ggplot(fas.df, aes(x=interval, y=mean.step, color=day)) +
   facet_grid(day~.) +
   geom_line() +
  labs(title = "Avg no of steps taken vs 5-min interval on Weekdays or Weekend", y="Avg no of steps", x= "5-min interval time")

plot.weekday.interval


```
